---
title: "RTSA SS23: Exercise Sheet 1"
output: html_document
author:
  - name: Martin Memmesheimer
  - name: Pranaya Tomar
date: "26.04.2023"
---

<!---------------- Some custom CSS. DO NOT CHANGE ANYTHING HERE --------------->
<style>

/*          this styles a box that we are going to use for feedback           */

.feedbackbox {
  padding: 1em;
  background: #042c58;
  color: white;
  border: 2px solid #042c58;  /* RPTU dunkelblau */
  border-radius: 5px;
}

</style>

```{r setup, include=FALSE}
# always keep this code chunk at the top
# it sets the global chunk option that code *and* output are rendered
# remember, we need to see your code in order to grade it
knitr::opts_chunk$set(echo = TRUE)
```

::: {.feedbackbox}
**FEEDBACK**

Leave this as it is, we fill it out while grading.
:::

<!------------------------- now, your part starts ----------------------------->

# Exercise 1.1

```{r}
# some general code for the exercise goes here
# or maybe loading the data
```

## Part (a)

You discussion for part (a) goes here.

```{r}
simulate_dice <- function(N) {
  frequencies <- c(0, 0, 0, 0, 0, 0)
  # roll this die n times
  for (i in c(1:N)){
    x <- sample((1:6), 1) # returns 1 sample number from 1 to 6
    frequencies[x] <- frequencies[x] + 1
  }
  relativ_frequencies <- frequencies / N
  return(relativ_frequencies)
}
```

## Part (b)

```{r}
x_values <- seq(from = 10, to = 200, by = 10) # values of N (number of rolls)
y_values <- lapply(lapply(x_values, simulate_dice), "[[", 4)
# inner lapply returns a list of relative_frequencies vectors from each run [[0.5,0.5,0.5,0.4,0.3,0.3], [0.5,0.5,0.5,0.4,0.3,0.3],...]
# outer lapply loops over the list of freq vecs and [[ extracts the frequency of 4 from each
plot(x_values, y_values, type = "o", ylim = c(0, 1), xlab = "N", ylab = "relative frequency of 4")
```

## Part (c)

```{r}
x_values <- seq(from = 10, to = 200, by = 10)
y_values <- lapply(lapply(x_values, simulate_dice), "[[", 4)
y_values_2 <- lapply(lapply(x_values, simulate_dice), "[[", 4)
y_values_3 <- lapply(lapply(x_values, simulate_dice), "[[", 4)
y_values_4 <- lapply(lapply(x_values, simulate_dice), "[[", 4)

plot(x_values, y_values, type = "o", ylim = c(0, 1), xlab = "N", ylab = "relative frequency of 4")
abline(0.166,0,col='red') #just to show the convergence to theoretical prob 1/6 

#add points to existing plot
points(x_values, y_values_2)
lines(x_values, y_values_2)
points(x_values, y_values_3)
lines(x_values, y_values_3)
points(x_values, y_values_4)
lines(x_values, y_values_4)
```

You can observe that the estimates scatter less as N gets larger. Also for N = 200 the estimations are only slightly different to the theoretical probability of $\frac{1}{6}$. This is explained by the law of large numbers.

Law of large numbers -  the average of the results obtained from a large number of trials should be close to the expected value and tends to become closer to the expected value as more trials are performed